---
title: "Potomac River Safety Analysis Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: cosmo
    css: custom.css
---

```{r setup, include=FALSE}
library(flexdashboard)
library(ggplot2)
library(dplyr)
library(DT)
library(plotly)

# Try to load advanced packages
use_advanced_features <- TRUE
tryCatch({
  library(dataRetrieval)
  library(lubridate)
}, error = function(e) {
  use_advanced_features <<- FALSE
})

# Get Potomac data
get_potomac_data <- function() {
  if (!use_advanced_features) {
    base_flow <- sample(c(1800, 2200, 2600, 3000, 2400), 1)
    return(list(
      flow_cfs = base_flow + rnorm(1, 0, 200),
      trend_cfs = rnorm(1, 0, 75),
      data_source = "demo_mode",
      success = TRUE,
      timestamp = Sys.time()
    ))
  }
  
  tryCatch({
    current_data <- dataRetrieval::readNWISuv(
      siteNumbers = "01646500",
      parameterCd = "00060",
      startDate = Sys.Date() - 7,
      endDate = Sys.Date()
    )
    
    if(nrow(current_data) > 0) {
      flows <- current_data$X_00060_00000
      current_flow <- tail(flows, 1)
      recent_flows <- tail(flows, min(96, length(flows)))
      trend_cfs <- if(length(recent_flows) > 10) {
        (tail(recent_flows, 1) - head(recent_flows, 1)) / length(recent_flows) * 96
      } else { 0 }
      
      return(list(
        flow_cfs = current_flow,
        trend_cfs = trend_cfs,
        data_source = "USGS_live",
        success = TRUE,
        timestamp = Sys.time()
      ))
    }
  }, error = function(e) {
    # Fallback to demo
  })
  
  # Demo fallback
  base_flow <- sample(c(1800, 2200, 2600, 3000, 2400), 1)
  return(list(
    flow_cfs = base_flow + rnorm(1, 0, 200),
    trend_cfs = rnorm(1, 0, 75),
    data_source = "demo_fallback",
    success = TRUE,
    timestamp = Sys.time()
  ))
}

# Safety scoring
calculate_safety_score <- function(flow_cfs, trend_cfs = 0) {
  if (flow_cfs < 800) {
    flow_score <- 0
  } else if (flow_cfs < 1200) {
    flow_score <- 15
  } else if (flow_cfs < 2000) {
    flow_score <- 35
  } else if (flow_cfs < 3000) {
    flow_score <- 40
  } else if (flow_cfs < 4000) {
    flow_score <- 30
  } else {
    flow_score <- 10
  }
  
  if (trend_cfs < -200) {
    trend_score <- 10
  } else if (trend_cfs < -50) {
    trend_score <- 20
  } else if (trend_cfs < 50) {
    trend_score <- 30
  } else if (trend_cfs < 200) {
    trend_score <- 25
  } else {
    trend_score <- 15
  }
  
  current_month <- as.numeric(format(Sys.Date(), "%m"))
  if (current_month %in% c(12, 1, 2)) {
    seasonal_score <- 10
  } else if (current_month %in% c(3, 4, 5)) {
    seasonal_score <- 20
  } else if (current_month %in% c(6, 7, 8)) {
    seasonal_score <- 15
  } else {
    seasonal_score <- 18
  }
  
  experience_score <- 5
  total_score <- flow_score + trend_score + seasonal_score + experience_score
  
  list(
    total = min(total_score, 100),
    flow_score = flow_score,
    trend_score = trend_score,
    seasonal_score = seasonal_score,
    experience_score = experience_score
  )
}

# Generate forecast
generate_forecast <- function(base_flow, base_trend = 0) {
  flows <- numeric(7)
  flows[1] <- base_flow
  
  for(i in 2:7) {
    trend_component <- base_trend * 0.1
    seasonal_component <- sin(2 * pi * i / 365) * 30
    mean_reversion <- (2000 - flows[i-1]) * 0.02
    random_component <- rnorm(1, mean = mean_reversion, sd = 150)
    change <- trend_component + seasonal_component + random_component
    flows[i] <- max(500, flows[i-1] + change)
  }
  
  scores <- numeric(7)
  risks <- character(7)
  
  for(i in 1:7) {
    trend_for_day <- if(i == 1) base_trend else (flows[i] - flows[max(1, i-1)]) * 96
    safety_result <- calculate_safety_score(flows[i], trend_for_day)
    scores[i] <- safety_result$total
    
    if (scores[i] >= 80) {
      risks[i] <- "EXCELLENT"
    } else if (scores[i] >= 65) {
      risks[i] <- "GOOD"
    } else if (scores[i] >= 45) {
      risks[i] <- "MODERATE"
    } else {
      risks[i] <- "POOR"
    }
  }
  
  data.frame(
    day = 1:7,
    predicted_flow = round(flows, 0),
    safety_score = round(scores, 0),
    risk_level = risks,
    trend_indicator = c("Current", rep("Forecast", 6))
  )
}

# Get current analysis
data_result <- get_potomac_data()
safety_metrics <- calculate_safety_score(data_result$flow_cfs, data_result$trend_cfs)

if (safety_metrics$total >= 80) {
  risk_level <- "EXCELLENT"
} else if (safety_metrics$total >= 65) {
  risk_level <- "GOOD"
} else if (safety_metrics$total >= 45) {
  risk_level <- "MODERATE"
} else {
  risk_level <- "POOR"
}

current_conditions <- list(
  flow_cfs = round(data_result$flow_cfs, 0),
  safety_score = safety_metrics$total,
  risk_level = risk_level,
  trend_cfs = round(data_result$trend_cfs, 1),
  timestamp = data_result$timestamp
)

forecast_data <- generate_forecast(data_result$flow_cfs, data_result$trend_cfs)
```

# Potomac River Safety Dashboard {.sidebar}

**🌊 Little Falls Analysis**

**Current Conditions:**
- **Flow:** `r format(current_conditions$flow_cfs, big.mark = ",")` cfs
- **Safety Score:** `r current_conditions$safety_score`/100
- **Risk Level:** `r current_conditions$risk_level`
- **Data Source:** `r if(data_result$data_source == "USGS_live") "🟢 USGS Live" else "🟡 Demo Mode"`

**Last Updated:** `r format(current_conditions$timestamp, "%Y-%m-%d %H:%M")`

---

**🔗 Part of bubulkaanalytics.com**  
*Analytics Dashboard Portfolio*

Column {data-width=600}
-----------------------------------------------------------------------

### 📈 7-Day Flow & Safety Forecast

```{r forecast_plot}
flow_scale <- max(forecast_data$predicted_flow) / 100

p <- ggplot(forecast_data, aes(x = day)) +
  geom_line(aes(y = predicted_flow), color = "#667eea", size = 1.5, alpha = 0.9) +
  geom_point(aes(y = predicted_flow), color = "#667eea", size = 3.5, alpha = 0.9) +
  geom_area(aes(y = safety_score * flow_scale), fill = "#764ba2", alpha = 0.25) +
  geom_line(aes(y = safety_score * flow_scale), color = "#764ba2", size = 1.5, alpha = 0.9) +
  geom_point(aes(y = safety_score * flow_scale), color = "#764ba2", size = 3.5, alpha = 0.9) +
  geom_text(aes(y = predicted_flow + max(predicted_flow) * 0.06, 
               label = substr(risk_level, 1, 1)),
           color = ifelse(forecast_data$risk_level == "EXCELLENT", "#27ae60",
                         ifelse(forecast_data$risk_level == "GOOD", "#f39c12",
                               ifelse(forecast_data$risk_level == "MODERATE", "#e67e22", "#e74c3c"))),
           size = 4.5, fontface = "bold") +
  scale_y_continuous(
    name = "Flow Rate (cfs)",
    labels = function(x) format(x, big.mark = ","),
    sec.axis = sec_axis(~ . / flow_scale, name = "Safety Score", breaks = seq(0, 100, 25))
  ) +
  labs(
    title = "7-Day Flow Rate & Safety Score Forecast",
    subtitle = paste("Current:", format(current_conditions$flow_cfs, big.mark = ","), 
                    "cfs | Score:", current_conditions$safety_score, 
                    "| Risk:", current_conditions$risk_level),
    x = "Days Ahead"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(color = "#2c3e50", face = "bold", size = 16),
    plot.subtitle = element_text(color = "#495057", size = 12),
    axis.text = element_text(color = "#495057"),
    axis.title = element_text(color = "#495057", face = "bold"),
    axis.title.y.right = element_text(color = "#764ba2"),
    axis.title.y.left = element_text(color = "#667eea"),
    panel.background = element_rect(fill = "white", color = NA),
    panel.grid.minor = element_blank(),
    legend.position = "none"
  ) +
  geom_hline(yintercept = c(800, 2000, 3000), linetype = "dashed", alpha = 0.4, color = "#95a5a6")

ggplotly(p, tooltip = c("x", "y"))
```

### 📋 7-Day Forecast Table

```{r forecast_table}
forecast_display <- forecast_data
forecast_display$Day <- ifelse(forecast_display$day == 1, "Today",
                              ifelse(forecast_display$day == 2, "Tomorrow",
                                     paste("Day", forecast_display$day)))
forecast_display$`Flow Rate` <- paste0(format(forecast_display$predicted_flow, big.mark = ","), " cfs")
forecast_display$`Safety Score` <- paste0(forecast_display$safety_score, "/100")

DT::datatable(
  forecast_display[, c("Day", "Flow Rate", "Safety Score", "risk_level", "trend_indicator")],
  colnames = c("Day", "Flow Rate", "Safety Score", "Risk Level", "Status"),
  options = list(pageLength = 7, dom = 't', ordering = FALSE),
  rownames = FALSE
) %>%
  DT::formatStyle(
    "risk_level",
    backgroundColor = DT::styleEqual(
      c("EXCELLENT", "GOOD", "MODERATE", "POOR"),
      c("#d4edda", "#fff3cd", "#f8d7da", "#f5c6cb")
    )
  )
```

Column {data-width=400}
-----------------------------------------------------------------------

### ⚖️ Safety Component Breakdown

```{r components_plot}
components_data <- data.frame(
  Component = c("Flow\nSafety", "Trend\nStability", "Seasonal\nFactor", "Experience\nBonus"),
  Score = c(safety_metrics$flow_score, safety_metrics$trend_score, 
           safety_metrics$seasonal_score, safety_metrics$experience_score),
  Max = c(40, 30, 20, 10)
)

components_data$Percentage <- (components_data$Score / components_data$Max) * 100
components_data$Color <- ifelse(components_data$Percentage >= 80, "#27ae60",
                               ifelse(components_data$Percentage >= 60, "#f39c12",
                                     ifelse(components_data$Percentage >= 40, "#e67e22", "#e74c3c")))

p2 <- ggplot(components_data, aes(x = reorder(Component, Score), y = Score, fill = Color)) +
  geom_col(width = 0.7, alpha = 0.8) +
  geom_text(aes(label = paste0(Score, "/", Max, "\n(", round(Percentage, 0), "%)")), 
            hjust = -0.1, color = "#2c3e50", fontface = "bold", size = 3.2) +
  coord_flip() +
  labs(title = "Safety Component Breakdown",
       subtitle = paste("Total Score:", sum(components_data$Score), "/ 100"),
       x = NULL, y = "Points Awarded") +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(color = "#2c3e50", face = "bold", size = 14),
    plot.subtitle = element_text(color = "#495057", size = 11),
    panel.background = element_rect(fill = "white", color = NA),
    legend.position = "none"
  ) +
  scale_fill_identity() +
  ylim(0, max(components_data$Max) * 1.4)

ggplotly(p2)
```

### 📊 Current Assessment

**Risk Assessment:**
```{r risk_text}
risk_text <- if (current_conditions$risk_level == "EXCELLENT") {
  "Ideal conditions for kayaking. All skill levels appropriate."
} else if (current_conditions$risk_level == "GOOD") {
  "Good conditions with minor considerations. Suitable for most paddlers."
} else if (current_conditions$risk_level == "MODERATE") {
  "Moderate conditions requiring attention. Intermediate+ skills recommended."
} else {
  "Poor conditions. Advanced skills required or consider postponing."
}
```

`r risk_text`

**Recommendations:**
```{r recommendations}
flow <- current_conditions$flow_cfs
trend <- current_conditions$trend_cfs

rec_text <- if (flow < 1200) {
  "Low water conditions - Watch for exposed rocks."
} else if (flow > 3500) {
  "High water conditions - Expect strong currents."
} else {
  "Flow within normal recreational range."
}

if (abs(trend) > 100) {
  rec_text <- paste(rec_text, "Rapidly changing conditions - Monitor closely.")
}
```

`r rec_text`

**System Status:**
- Data Source: `r if(data_result$data_source == "USGS_live") "USGS Live Feed" else "Demo Mode"`
- Station ID: 01646500
- Last Update: `r format(current_conditions$timestamp, "%H:%M:%S")`